In mammals, an important proportion of advantageous mutations are repairing damaged genomes instead of creating adaptive innovations

### 2022-10-26 manuscript/main_rev.tex
Revision by L. Duret & J. Joseph

\usepackage{biorxiv}
\usepackage[backend=biber,,isbn=false,url=false,intitle=true,style=nature]{biblatex}
\addbibresource{codon_models.bib}

@@

    \begin{abstract}
        Positive selection can be the result of adaptive evolution, driven by a change in environment or selective pressure.
        However, positive selection can also be the result of what we call maintenance positive selection where advantageous mutations are reverting deleterious substitutions without any change of the environment.
        The evolutionary outcomes of these two positive selections are drastically different.
        While adaptive evolution acts as a diversifying force, maintenance positive selection stabilizes phenotypes and reduces diversity across species.
        By leveraging polymorphism data in 28 populations and divergence data between 94 mammalian species, we estimate that around 25\% of observed substitutions are owed to maintenance positive selection.
        Our work confirms that deleterious substitutions have accumulated in mammals and are currently being eliminated, resulting in widespread positive selection that is not adaptive evolution.
    \end{abstract}

@@

    %\NS{on devrait plus placer le départ de l'intro dans un contexte historique, pour dire que de connaître si une mutation est avantageuse est la base de la génétique moléculaire (en citant des papiers clés)}
    Adaptation is considered as the main process driving the diversity of forms and functions in the tree of life\cite{darwin_origin_1859}.
    It allows species to live in new ecological niches, and to evolve in response to a change in environment\cite{darwin_origin_1859, merrell_adaptive_1994}.
    The notion of adaptation, in its eco-evolutionary sense, is tightly linked to the notion of change, both of the environment (biotic and abiotic), and of the species responding to the latter\cite{merrell_adaptive_1994}.
    With the discovery of DNA, the main support of heritable information, evolutionary geneticists looked for signatures of adaptation in genomes, such as to quantify it\cite{mcdonald_adaptative_1991, eyre-walker_distribution_2006}.
    In that context arose the term of rate of adaptive evolution\cite{mcdonald_adaptative_1991, eyre-walker_distribution_2006, galtier_adaptive_2016}.
    However, in contrast to adaptation in its eco-evolutionary sense, adaptive evolution is merely used as a synonymous of evolution under positive selection\cite{eyre-walker_distribution_2006}.
    This leads to a misunderstanding on what is adaptation with the loss of the notion of change, and a misinterpretation of the so-called rate of adaptive evolution.
    Indeed, adaptive evolution is not the only process which can lead to positive selection at the population scale\cite{sella_application_2005, mustonen_fitness_2009}.
    At mutation-selection-drift equilibrium, a mutation restoring the ancestral allele which corresponded to an optimal fitness in a constant environment, but lost due to the fixation of a deleterious mutation by drift, will also be positively selected (fig.~\ref{fig:fitness-landscape}B).
    In this study, this process will be referred to as maintenance positive selection.
    %At the scale of a population, adaptive evolution and maintenance positive selection display different dynamics.
    %While in the case of environmental change, the population as a whole is climbing the fitness landscape along an adaptive walk\cite{tenaillon_utility_2014}, in the case of maintenance positive selection, the population as a whole is at mutation-selection equilibrium and therefore merely stays at the same location in the fitness landscape\cite{sella_application_2005, mustonen_fitness_2009}.
    %We can distinguish two main scenarios that can lead to such positive selection.
    %In the first scenario, a mutation towards a state which was deleterious is now advantageous because of a change of environment.
    %A change of environment could be a change of temperature, viral exposition but it could also be a change in another part of the species' genome.
    %In the new environment, the mutation allows its bearer to survive and reproduce more than the others and thus will be positively selected.
    %This process is typically referred to as adaptive evolution (fig.~\ref{fig:fitness-landscape}A), and meets the eco-evolutionary definition of adaptation.
    At the population scale, maintenance positive selection and adaptive evolution are indistinguishable and both result in a positive transmission bias of the advantageous allele, but their consequences at a macro-evolutionary scale are fundamentally different.
    While adaptive evolution promotes phenotype diversification (fig.~\ref{fig:fitness-landscape}C), maintenance positive selection promotes phenotype stability and preserves well established biological systems (fig.~\ref{fig:fitness-landscape}D).
    It is thus essential to distinguish these two types of positive selection to understand the long term effects of selection on the diversification of species.

    At the gene level, positive selection can be quantified by comparing sequence variation inside a population with the sequence divergence from a sister species\cite{mcdonald_adaptative_1991}.
    The basic idea is that advantageous mutations are expected to be observed in substitutions between species but not in polymorphism within a population because they reach fixation too quickly.
    This method has been largely improved to account for possible bias-inducing factors such as the presence of weak positive selection, background selection and demographic fluctuations\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    It has been applied across many organisms and enabled the discovery of widespread positive selection\cite{moutinho_variation_2019}.
    However, the exact causes for this widespread occurrence of positive selection have been scarcely investigated\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016} %\NS{is it the case in the conclusion of the previous paper?? would be nice to back it up with references}.
    As mentioned above, not all advantageous mutations contribute to genetic innovation and diversification, and truly adaptive evolution can only be well estimated if maintenance positive selection is taken into account.
    While adaptive evolution is unpredictable because it caused by an unpredictable change in the environment of organisms, the second is predictable because optimal states should be more or less the same for closely related organisms which share most of their environment and physiology.
    A way to thus measure maintenance positive selection is to predict selection coefficients assuming a constant and stable fitness landscape estimated from the alignment of protein-coding genes.
    For each site of the protein alignment, a first approximation would be to consider that the frequency of an amino acid is a proxy for its fitness, since for example a deleterious amino acid will be more likely to be absent in the alignment.
    One of the most used methods to make these predictions is SIFT (Sort Intolerant From Tolerant), which reliably detects deleterious mutations in many taxa\cite{ng_sift_2003, vaser_sift_2016}.
    However, this method has two major shortcomings.
    First, we know that a lot of mechanisms can influence amino acid frequency other than fitness (phylogeny, mutation biases, GC-biased gene conversion, mutational paths) which are not accounted for in SIFT\@.
    Second, SIFT scores is not interpretable as a selection coefficient.
    Contrarily, so-called mutation-selection codon models are rooted within a population genetics formalism\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    These phylogenetic codon models can disentangle processes of mutation at the DNA level and selection at the amino acid level.
    Such codon models provide a nearly-neutral framework of evolution by estimating the fitness landscape over amino acid sequences for each site of the sequence (fig~\ref{fig:method}A) while taking into account non-adaptive processes impacting codon frequency\cite{halpern_evolutionary_1998, rodrigue_mechanistic_2010, tamuri_estimating_2012, latrille_improved_2022}.
    From the fitness landscape obtained by fitting the model to the data, one can thus compute the scaled selection coefficient ($\Sphy$) of any non-synonymous mutations.
    Importantly, mutation-selection codon models assume that the underlying fitness landscape is stable along the phylogeny (as in fig.~\ref{fig:fitness-landscape}B), thus a mutation with a positive $\Sphy$ would be driving the current site toward a fitter amino-acid, and such mutation is effectively considered under maintenance positive selection.

    By integrating datasets at the species and population level, we can predict $\Sphy$ for each mutation currently segregating in a population (fig~\ref{fig:method}B).
    Such a dataset can allow us to address different questions regarding positive selection affecting protein-coding DNA sequences.
    Which proportion of the segregating polymorphisms in population are predicted toward a more optimal amino acid (positive $\Sphy$), thus supposedly under maintenance positive selection?
    Are these polymorphisms effectively under positive selection at the scale of populations?
    Finally, how much of positive selection observed in populations is due to maintenance positive selection?
    To answer these questions, we estimated the fitness landscape of amino-acids for long-term evolution using available multiple sequence alignments between mammalian species\cite{ranwez_orthomam_2007, howe_ensembl_2021}.
    We then integrated data at the phylogenetic and population scale to estimate divergence and polymorphism, respectively, across the entire exome for 28 populations from 6 genera (\textit{Equus},  \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}; fig.~\ref{fig:method}A\&B).

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method}
        \caption{
            Integrating divergence and polymorphism.
            At the phylogenetic level (panel A), the amino acid Wrightian fitness for each site is estimated from protein-coding DNA alignments using mutation-selection codon models.
            At the population-genetic level (panel B), for each observed single nucleotide polymorphism (SNP) segregating in the population, selection coefficient ($\Sphy$) are computed as the difference in amino acid fitness between the ancestral and derived variants.
            Mutation-selection codon models that the underlying fitness landscape is stable along the phylogeny, taking into account only non-adaptive evolution, thus a mutation with a positive $\Sphy$ would be driving the current site toward are fitter amino-acid, and such mutation are effectively considered under maintenance positive selection.
        }
        \label{fig:method}
    \end{figure*}

@@

    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) from amino acid fitnesses estimated at the mammalian scale (fig.~\ref{fig:method}B).
    Then we ask whether the variants that are predicted to be advantageous, thus supposedly under maintenance selection, are showing signatures of short term positive selection.
    For example, if a variant is found to be fixed in many mammals but not in ancestral humans, and a mutation in humans is toward the mammalian version, is this mutation advantageous for the individual carrying it?
    At the population level (e.g.~in humans of Asian descent in fig.~\ref{fig:homo-afr-results}), we divided all observed SNPs in 5 classes of predicted selection coefficient: severely deleterious with $\divStrongDel$, deleterious with $\divDel$, weakly deleterious with $\divWeakDel$, weakly advantageous with $\divWeakAdv$ and advantageous with $\divAdv$ (fig.~\ref{fig:homo-afr-results}B).
    In each of theses classes, SNPs are then grouped by the number of derived alleles in the population (fig.~\ref{fig:homo-afr-results}C), generating a so called site-frequency spectrum (SFS).
    Compared to synonymous mutations which are supposedly neutral (black line in fig.~\ref{fig:homo-afr-results}C), mutations predicted to be deleterious (blue and green lines in fig.~\ref{fig:homo-afr-results}C) are underrepresented at high frequencies in the population, a signature of ongoing purifying selection.
    However, these observations remain qualitative.

@@

    We replicated such experiment across 28 populations (fig.~\ref{fig:heatmap}), and found that mutations predicted to be deleterious are indeed purged at the population-genetic scale (blue and green rows in fig.~\ref{fig:heatmap}A).
    However, mutations predicted to be weakly advantageous ($\divWeakAdv$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (orange row in fig.~\ref{fig:heatmap}B).
    Similarly, mutations predicted to be advantageous ($\divAdv$, under maintenance selection) are found to be positively selected with varying proportions across populations (red row in fig.~\ref{fig:heatmap}C).
    In other words, mutations predicted to be deleterious by the mutation-selection model are effectively purified away in any population, while mutations predicted to be advantageous can be either neutral or advantageous depending on the population.
    These variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, a reasonable proxy usually used is the within a population (Watterson $\theta_w$).
    We thus ordered populations by their levels of synonymous diversity (fig.~\ref{fig:diversity}A) and confirmed that higher synonymous diversity is typically accompanied by a smaller proportion of neutral mutations.
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate between advantageous and deleterious mutations (fig~S1-S5).
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.
    %, especially for mutations already predicted to be nearly-neutral at the phylogenetic scale (fig~S8-S9).

    Finally, we reconstructed the substitutions in the terminal branches of the phylogenetic tree to determine the distribution of their selection coefficients.
    For each population with available polymorphism data, we first reconstructed the ancestral sequence (which is different than the reference sequence), then we mapped the substitutions from the last genus split to this reconstructed ancestral sequence.
    In these populations, we show that out of all polarized substitutions in the branches, between 10 and 13\% of all substitutions have an $\Sphy > 1$ (table S1) and between 25 and 30\% of all substitutions have an $\Sphy > 0$ (table S3).
    Most importantly, there is clear signal that the substitutions with an $\Sphy > 1$ have been positively selected since their ratio of non-synonymous over synonymous divergence, a ratio called $\dnds$, is between 1.15 and 1.68 in the different populations (table S1), meaning that mutations under maintenance selection with $\Sphy > 1$ reach fixation more frequently that synonymous mutations (supposedly neutral).
    These observed substitutions are likely to have been positively selected because they restored a more optimal state, not because they are driving innovations to a new environment or a new function.
    %This also means that the ratio $\dnds$ usually used to detect adaptation while relying on substitutions between close species\cite{mcdonald_adaptative_1991, galtier_adaptive_2016} or at the phylogenetic scale\cite{goldman_codonbased_1994, yang_codonsubstitution_2002} is overestimated.

@@

    We then sought to estimate the proportion of positive selection that is due to maintenance selection.
    Mathematically, this corresponds to $\mathbb{P} [ \Sphy > 1  \given  \Spop > 1]$, namely the probability for a mutation to be due to maintenance selection, given it is positively selected.
    This probability is given by Bayes formula (see methods), given we have access to $\mathbb{P} [ \Sphy > 1 ]$, $\mathbb{P} [ \Spop > 1 ]$, $\mathbb{P} [ \Spop > 1  \given  \Sphy > 1]$.
    First, $\mathbb{P} [ \Sphy > 1 ]$ the probability for a mutation away from the ancestral sequence to have $\Sphy > 1$ (red fill of histogram in fig~\ref{fig:homo-afr-results}A and table S2).
    Second, $\mathbb{P} [ \Spop > 1 ]$ is is the part of the DFE estimated at the population-scale greater than 1 when we used all the SNPs available in the population (black fill in fig~\ref{fig:diversity}A and table S2).
    Finally, $\mathbb{P} [ \Spop > 1  \given  \Sphy > 1]$ is the part of the DFE estimated at the population-scale greater than 1, when we restricted our analysis on SNPs that are already predicted by mutation-selection models to have a selection coefficient greater than 1 (black fill for the category $\Sphy > 1$ in fig~\ref{fig:homo-afr-results}D and table S2).
    Altogether, the probability for a mutation to be restoring the genome given that it is positively selected is obtained for each population, and varies between 10 and 26\% (table S2).
    % Since adaptation does not fit the fitness predictions, the proportion of adaptive mutation is expected to be the same among SNPs which are predicted to be advantageous, and among SNPs which are predicted to be deleterious.
    % Thus the proportion of advantageous mutations among SNPs which are predicted to be deleterious directly correspond to the proportion of all mutations which are adaptive.
    % So if we divide the proportion of mutations which are predicted to be advantageous among supposedly deleterious SNPs by the proportion of mutations which are predicted to be advantageous among all SNPs, we directly get the proportion of mutations which are adaptive among advantageous ones.
    % It results that 20 to 40\% of advantageous mutations are due to maintenance positive selection (fig~\ref{fig:diversity}B \& table S2).

    Focusing on substitutions in the terminal lineages, we can consider that the total $\dnds$ encompasses the rate of evolution for deleterious and advantageous (both adaptive and non-adaptive) substitutions.
    Then, the $\dnds$ computed exclusively on non-synonymous sites with a $\Sphy < 0$ only represent the rate of evolution of solely deleterious and adaptive substitutions, discarding maintenance selection by construction.
    Thus, if we divide the $\dnds$ estimated on all substitutions, by the one estimated only on substitutions predicted to be deleterious, we directly get the proportion of the rate of evolution which is contributed by maintenance selection (see methods).
    This way, we estimated that between 22 and 27\% of the evolutionary rate is due to maintenance positive selection (fig~\ref{fig:diversity}C \& table S3).

>>>>

\usepackage{biorxiv}
\usepackage[backend=biber,eprint=false,isbn=false,url=false,intitle=true,style=nature,date=year]{biblatex}
\addbibresource{codon_models.bib}

@@

    \begin{abstract}
        Positive selection can be the result of adaptive evolution, driven by a change in environment or selective pressure.
        However, positive selection can also be the result of what we call maintenance positive selection where advantageous back-mutations are reverting deleterious substitutions without any change of the environment.
        The evolutionary outcomes of these two positive selections are drastically different.
        While adaptive evolution acts as a diversifying force, maintenance positive selection stabilizes phenotypes and reduces diversity across species.
        By leveraging polymorphism data in 28 populations and divergence data between 94 mammalian species, we estimate that around 25\% of observed substitutions are owed to maintenance positive selection.
        Our work confirms that deleterious substitutions have accumulated in mammals and are currently being eliminated, resulting in widespread positive selection that is not adaptive evolution.
    \end{abstract}

@@

    Adaptation is considered as the main process driving the diversity of forms and functions across the tree of life\cite{darwin_origin_1859}, and it enables species to live in new ecological niches, and to evolve in response to a change in environment\cite{darwin_origin_1859, merrell_adaptive_1994}.
    The notion of adaptation, in its eco-evolutionary sense, is tightly linked to the notion of change, both of the environment (biotic and abiotic), and of the species responding to the latter\cite{merrell_adaptive_1994}.
    With the discovery of DNA, the main support of heritable information, evolutionary geneticists looked for signatures of adaptation in genomes, and tried to quantify it\cite{mcdonald_adaptative_1991, eyre-walker_distribution_2006}.
    In that context arose the term of rate of adaptive evolution\cite{mcdonald_adaptative_1991, eyre-walker_distribution_2006, galtier_adaptive_2016}.
    However, in contrast to adaptation in its eco-evolutionary sense, adaptive evolution is merely used as a synonymous of evolution under positive selection\cite{eyre-walker_distribution_2006}.
    This leads to a misunderstanding on what is adaptation with the loss of the notion of change, and a misinterpretation of the so-called rate of adaptive evolution.
    Indeed, adaptive evolution is not the only process which can lead to positive selection at the population scale\cite{sella_application_2005, mustonen_fitness_2009}.
    At mutation-selection-drift equilibrium, a mutation restoring the ancestral allele which corresponded to an optimal fitness in a constant environment, but lost due to the fixation of a deleterious mutation by drift, will also be positively selected (fig.~\ref{fig:fitness-landscape}B).
    In this study, this process will be referred to as maintenance positive selection.
    At the population scale, maintenance positive selection and adaptive evolution are indistinguishable and both result in a positive transmission bias of the advantageous allele, but their consequences at a macro-evolutionary scale are fundamentally different.
    While adaptive evolution promotes phenotype diversification (fig.~\ref{fig:fitness-landscape}C), maintenance positive selection promotes phenotype stability and preserves well established biological systems (fig.~\ref{fig:fitness-landscape}D).
    It is thus essential to distinguish these two types of positive selection to understand the long term effects of selection on the diversification of species.

    At the gene level, positive selection can be quantified by comparing sequence variation inside a population with the sequence divergence from a sister species\cite{mcdonald_adaptative_1991}.
    The basic idea is that advantageous mutations are expected to be observed in substitutions between species but not in polymorphism within a population because they reach fixation too quickly.
    This method has been largely improved to account for possible bias-inducing factors such as the presence of weak positive selection, background selection and demographic fluctuations\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016, tataru_inference_2017}.
    It has been applied across many organisms and enabled the discovery of widespread positive selection\cite{moutinho_variation_2019}.
    However, the exact causes for this widespread occurrence of positive selection have been scarcely investigated\cite{eyre-walker_distribution_2006, eyre-walker_estimating_2009, galtier_adaptive_2016} %\NS{is it the case in the conclusion of the previous paper?? would be nice to back it up with references}.
    As mentioned above, not all advantageous mutations contribute to genetic innovation and diversification, and truly adaptive evolution can only be well estimated if maintenance positive selection is taken into account.
    While adaptive evolution is unpredictable because it caused by an unpredictable change in the environment of organisms, the second is predictable because optimal states should be more or less the same for closely related organisms which share most of their environment and physiology.
    A way to measure maintenance positive selection is to predict selection coefficients assuming a constant and stable fitness landscape estimated from the alignment of protein-coding genes.
    For each site of the protein alignment, a first approximation would be to consider that the frequency of an amino acid is a proxy for its fitness, since for example a deleterious amino acid will be more likely to be absent in the alignment.
    One of the most used methods to make these predictions is SIFT (Sort Intolerant From Tolerant), which reliably detects deleterious mutations in many taxa\cite{ng_sift_2003, vaser_sift_2016}.
    However, this method has two major shortcomings.
    First, we know that a lot of mechanisms can influence amino acid frequency other than fitness (phylogenetic inertia, mutation biases, GC-biased gene conversion, mutational paths) which are not accounted for in SIFT\@.
    Second, SIFT scores is not interpretable as a selection coefficient.
    Contrarily, so-called mutation-selection codon models are rooted within a population genetics formalism\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    These phylogenetic codon models can disentangle processes of mutation at the DNA level and selection at the amino acid level.
    Such codon models provide a nearly-neutral framework of evolution by estimating the fitness landscape over amino acid sequences for each site of the sequence (fig~\ref{fig:method}A) while taking into account non-adaptive processes impacting codon frequency\cite{halpern_evolutionary_1998, rodrigue_mechanistic_2010, tamuri_estimating_2012, latrille_improved_2022}.
    From the fitness landscape obtained by fitting the model to the data, one can thus compute the scaled selection coefficient ($\Sphy$) of any non-synonymous mutations.
    Importantly, mutation-selection codon models assume that the underlying fitness landscape is stable along the phylogeny (as in fig.~\ref{fig:fitness-landscape}B), thus a mutation with a positive $\Sphy$ would be driving the current site toward a fitter amino-acid, and such mutation is effectively considered under maintenance positive selection, and is non-adaptive.

    By integrating datasets at the species and population level, we can predict $\Sphy$ for any mutation currently segregating in a population (fig~\ref{fig:method}B) under a stable fitness landscape.
    Such a dataset can allow us to address different questions regarding positive selection affecting protein-coding DNA sequences.
    Which proportion of the segregating polymorphisms in population are predicted toward a more optimal amino acid (positive $\Sphy$), thus supposedly under maintenance positive selection?
    Are these polymorphisms effectively under positive selection along the terminal branches ? At the scale of populations ?
    Finally, how much of positive selection observed in populations is due to maintenance positive selection?
    To answer these questions, we estimated the fitness landscape of amino-acids for long-term evolution using available multiple sequence alignments between mammalian species\cite{ranwez_orthomam_2007, howe_ensembl_2021}.
    We then integrated data at the phylogenetic and population scale to estimate divergence and polymorphism, respectively, across the entire exome for 28 populations from 6 genera (\textit{Equus},  \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus} and \textit{Homo}; fig.~\ref{fig:method}A\&B).

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method}
        \caption{
            Integrating divergence and polymorphism.
            At the phylogenetic level (panel A), the amino acid Wrightian fitness for each site is estimated from protein-coding DNA alignments using mutation-selection codon models.
            At the population-genetic level (panel B), for each observed single nucleotide polymorphism (SNP) segregating in the population, selection coefficient ($\Sphy$) are estimated as the difference in amino acid fitness between the ancestral and derived variants.
            Mutation-selection codon models assume that the underlying fitness landscape is stable along the phylogeny, taking into account only non-adaptive evolution, thus a mutation with a positive $\Sphy$ would be driving the current site toward are fitter amino-acid, and such mutation are effectively considered under maintenance positive selection.
        }
        \label{fig:method}
    \end{figure*}

@@

    For each non-synonymous SNP observed at the population level, we predicted its scaled selection coefficient ($\Sphy$) from a stable fitness landscape estimated at the mammalian scale (fig.~\ref{fig:method}B).
    Then we ask whether the variants that are predicted to be advantageous, thus supposedly under maintenance selection, are showing signatures of short term positive selection.
    For example, if a variant predicted to be non-optimal is found to be fixed in the last common ancestor of extant humans. If a mutation occurs in extant humans towards the allele predicted to be optimal, is this mutation positively selected at the population-scale?
    At the population level (e.g.~in humans of Asian descent in fig.~\ref{fig:homo-afr-results}), we divided all observed SNPs in 5 classes of predicted selection coefficient: severely deleterious with $\divStrongDel$, deleterious with $\divDel$, weakly deleterious with $\divWeakDel$, weakly advantageous with $\divWeakAdv$ and advantageous with $\divAdv$ (fig.~\ref{fig:homo-afr-results}B).
    In each of theses classes, SNPs are then grouped by the frequency of derived alleles in the population (fig.~\ref{fig:homo-afr-results}C), generating a so called site-frequency spectrum (SFS).
    Compared to synonymous mutations which are supposedly neutral (black line in fig.~\ref{fig:homo-afr-results}C), mutations predicted to be deleterious (blue and green lines in fig.~\ref{fig:homo-afr-results}C) are underrepresented at high frequencies in the population, a signature of ongoing purifying selection.
    However, these observations remain qualitative.

@@

    We replicated such experiment across 28 populations from 6 species(fig.~\ref{fig:heatmap}), and found that mutations predicted to be deleterious are indeed purged at the population-scale (blue and green rows in fig.~\ref{fig:heatmap}A).
    However, mutations predicted to be weakly advantageous ($\divWeakAdv$) are composed of a mix of neutral and selected mutations with varying proportions across the different populations (orange row in fig.~\ref{fig:heatmap}B).
    Similarly, mutations predicted to be advantageous ($\divAdv$, under maintenance selection) are found to be positively selected with varying proportions across populations (red row in fig.~\ref{fig:heatmap}C).
    In other words, mutations predicted to be deleterious by the mutation-selection model are effectively purified away in any population, while mutations predicted to be advantageous can be either neutral or advantageous depending on the population.
    These variable proportions between populations can be explained by the effective number of individuals in the population ($\Ne$), a major driver of selection efficacy.
    Since $\Ne$ is not directly measurable, a reasonable proxy usually used is the neutral diversity (Watterson $\theta_w$).
    Assuming that synonymous positions are effectively neutral in mammals, we thus ordered populations by their levels of synonymous diversity (fig.~\ref{fig:diversity}A) and confirmed that higher synonymous diversity is typically accompanied by a smaller proportion of neutral mutations.
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate between advantageous and deleterious mutations (fig~S1-S5).
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.
    %, especially for mutations already predicted to be nearly-neutral at the phylogenetic scale (fig~S8-S9).

    Finally, we reconstructed the substitutions in the terminal branches of the phylogenetic tree to determine the distribution of their selection coefficients.
    For each population with available polymorphism data, we first reconstructed the ancestral sequence by replacing all variable site of the reference genome with the ancestral allele predicted by est-usfs, then we mapped the substitutions from the last genus split to this reconstructed ancestral sequence.
    In these populations, we show that out of all polarized substitutions in the branches, between 10\% and 13\% of all substitutions have an $\Sphy > 1$ (table S1) and between 25\% and 30\% of all substitutions have an $\Sphy > 0$ (table S3).
    Most importantly, there is clear signal that the substitutions with an $\Sphy > 1$ have been positively selected since their ratio of non-synonymous over synonymous divergence, a ratio called $\dnds$, is between 1.15 and 1.68 in the different populations (table S1), meaning that mutations under maintenance selection with $\Sphy > 1$ reach fixation more frequently that synonymous mutations (supposedly neutral).
    These observed substitutions are likely to have been positively selected because they restored a more optimal state, not because they are driving innovations to a new environment or a new function.
    %This also means that the ratio $\dnds$ usually used to detect adaptation while relying on substitutions between close species\cite{mcdonald_adaptative_1991, galtier_adaptive_2016} or at the phylogenetic scale\cite{goldman_codonbased_1994, yang_codonsubstitution_2002} is overestimated.

@@

    We then sought to estimate the proportion of positive selection that is due to maintenance selection.
    Mathematically, this corresponds to $\mathbb{P} [ \Sphy > 1  \given  \Spop > 1]$, namely the probability for a mutation to be due to maintenance selection, given it is positively selected.
    This probability is given by Bayes formula (see methods), given we have access to $\mathbb{P} [ \Sphy > 1 ]$, $\mathbb{P} [ \Spop > 1 ]$, $\mathbb{P} [ \Spop > 1  \given  \Sphy > 1]$.
    First, $\mathbb{P} [ \Sphy > 1 ]$ the probability for a mutation away from the ancestral sequence to have $\Sphy > 1$ (red fill of histogram in fig~\ref{fig:homo-afr-results}A and table S2).
    Second, $\mathbb{P} [ \Spop > 1 ]$ is is the part of the DFE estimated at the population-scale greater than 1 when we used all the SNPs available in the population (black fill in fig~\ref{fig:diversity}A and table S2).
    Finally, $\mathbb{P} [ \Spop > 1  \given  \Sphy > 1]$ is the part of the DFE estimated at the population-scale greater than 1, when we restricted our analysis on SNPs that are already predicted by mutation-selection models to have a selection coefficient greater than 1 (black fill for the category $\Sphy > 1$ in fig~\ref{fig:homo-afr-results}D and table S2).
    Altogether, the probability for a mutation to be restoring the genome given that it is positively selected is obtained for each population, and varies between 10\% and 26\% depending of the population (table S2).
    % Since adaptation does not fit the fitness predictions, the proportion of adaptive mutation is expected to be the same among SNPs which are predicted to be advantageous, and among SNPs which are predicted to be deleterious.
    % Thus the proportion of advantageous mutations among SNPs which are predicted to be deleterious directly correspond to the proportion of all mutations which are adaptive.
    % So if we divide the proportion of mutations which are predicted to be advantageous among supposedly deleterious SNPs by the proportion of mutations which are predicted to be advantageous among all SNPs, we directly get the proportion of mutations which are adaptive among advantageous ones.
    % It results that 20 to 40\% of advantageous mutations are due to maintenance positive selection (fig~\ref{fig:diversity}B \& table S2).

    Focusing on substitutions in the terminal lineages, we can consider that the total $\dnds$ encompasses the rate of evolution for deleterious and advantageous (both adaptive and non-adaptive) substitutions.
    Then, the $\dnds$ computed exclusively on non-synonymous sites with a $\Sphy < 0$ only represent the rate of evolution of solely deleterious and adaptive substitutions, discarding maintenance selection by construction.
    Thus, if we divide the $\dnds$ estimated on all substitutions, by the one estimated only on substitutions predicted to be deleterious, we directly get the proportion of the rate of evolution which is contributed by maintenance selection (see methods).
    This way, we estimated that between 22\% and 27\% of the evolutionary rate is due to maintenance positive selection (fig~\ref{fig:diversity}C \& table S3).
